---
title: "An Introduction to EETE"
author: "Michelle Kairy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Introduction to EETE}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r}
library(tidyverse)
library(haven)
library(eete)
```

```{r}
SIPP1991 = read_dta('../data/Angrist 1991 SIPP.dta')
SIPP1991_edit <- SIPP1991 |>
  mutate(rsncode = ifelse(rsncode == 999, NA, rsncode),
         wage = kwage,
         lwage = log(kwage))
```


This package provides essential functions to calculate the egalitarian equivalent treatment effect.

## Functions for Evaluators

### Constant Difference Protected Income Evaluations (CDPIE)

If given a value for $r$, *cdpie*, calculates $$g(x) = -e^{-\frac{x}{r}}$$ If given a value for $L$, *cdpie*, calculates $$g(x) = -2^{-\frac{x}{L}}$$

-   **x** A numeric value or list.
-   **r** The value of r, a scale factor that adjusts the intensity of risk aversion. This parameter is optional if L is provided.
-   **L** The value of L. This parameter is optional if r is provided.

**return** A numeric or list of return values from the CACD utility function.


### Constant Relative Protected Income Evaluations (CRPIE)

If given an value for $\gamma$, *crpie* calculates $$g(x) = \frac{x^{1-\gamma}}{1-\gamma}$$

Or, if given a value for $k$, *crpie* calculate $$g(x) = \log_2(k) x^{\frac{1}{\log_2(k)}}$$

*Note: If* $k < 1$, $k$ represents the percent of income to protect. If $k > 1$, $k$ represents the multiple person A must receive to sacrifice income of person B completely.

-   **x** A numeric value or list.
-   **gam** The value of gamma, the coefficient of relative risk aversion. This value is optional if a value for k is provided.
-   **k** The value of k, the percent of income to protect. This value is optional if a value for gamma is provided.

**return** A numeric or list of return values from the CRCD utility function.

## The **"eete()"** function

This function estimates the eete, which optional arguments to include an IV or bootstrapped SE. It applies an input function to data, conditioning on treatment (and instrument variables, if applicable).

-   **inputFunction** A function that will be applied to the data.
-   **...** Additional arguments to be passed to `inputFunction`.
-   **y** The name of the outcome variable column in `data`.
-   **d** The name of the treatment variable column in `data`.
-   **z** Optional argument to include the name of the instrument variable column in `data`.
-   **data** The data frame containing the variables.
-   **lower** The lower bound for finding the inverse function value.
-   **upper** The upper bound for finding the inverse function value.
-   **se** Optional argument to calculate the standard error of the estimate with bootstrapping.
-   **B** The number of iterations to be used in the standard error bootstrap.

**return** The estimated egalitarian equivalent treatment effect.

```{r}
eete(crpie, gam = 0, y = "lwage", d = "nvstat", z = "rsncode", data = SIPP1991_edit)
eete(cdpie, r = 1, y = "lwage", d = "nvstat", z = "rsncode", data = SIPP1991_edit)
```


## The **"ebounds()"** function

### Documentation

This function returns a data frame of the upper and lower bounds calculations.

-   **inputFunction** A function that will be applied to the data.
-   **...** Additional arguments to be passed to `inputFunction`.
-   **y** The name of the outcome variable column in `data`.
-   **d** The name of the treatment variable column in `data`.
-   **z** Optional argument to include the name of the instrument variable column in `data`.
-   **wgt** A vector of weights.
-   **data** The data frame containing the variables.
-   **se** Optional argument to calculate the standard error of the estimate with bootstrapping.
-   **B** The number of iterations to be used in the standard error bootstrap.

**return** A data frame with probabilities and bounds.

```{r}
ebounds(crpie, gam = 2, y = "lwage", d = "nvstat", z = "rsncode", data = SIPP1991_edit, se = TRUE)
```


